---
title: "Checking density values for Suarez et al 2018"
author: "Tati Micheletti"
date: "7/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why this file?

Talking to Nicole and Alberto to understand where the data from Alberto's "Final_points_2010.csv" 
table come from in order to be able to predict correctly using the same abundances, I came across
a situation where Alberto used a file named `DDB_BCR_PROV_LCC_Apr4-2012.csv` while Nicole suggested 
that the right densities should come from a file named `Habitat_Association_by_jurisdiction(bamddb_Apr19-2012).csv`.

Therefore, to analyze/understand how much the manuscript was affected by using a different file, I 
started the following check-up. First thing is to place both files in folder `~/inputs/`.

### Loading the tables
```{r load_tables}
usedDens <- data.table::fread(file.path(getwd(), "inputs", "DDB_BCR_PROV_LCC_Apr4-2012.csv"))
corrDens <- data.table::fread(file.path(getwd(), "inputs", "Habitat_Association_by_jurisdiction(bamddb_Apr19-2012).csv"))
usedDens
corrDens
```

### Testing for unique LCC and BCR values




