---
title: "Testing CONW dataset"
author: "Tati Micheletti"
date: "3/13/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup}
library(data.table)
library(lme4)
```

## Dataset

```{r table, include=FALSE}
datasetCONW <- data.table(readRDS(file.path(getwd(),"datasetCONW.rds")))
knitr::kable(datasetCONW)
```

### These models are not working:
```{r modelsNot, error=TRUE}
CONW = glmer(AB_CONW ~ State_P_500 + LOG_BCR_CONW + offset(OF_CONW) + (1|ClusterSP) + (1|YYYY) + (1|ClusterSP:YYYY), family="poisson", data=datasetCONW)
CONW.noIntRE = glmer(AB_CONW ~ State_P_500 + LOG_BCR_CONW + offset(OF_CONW) + (1|ClusterSP) + (1|YYYY), family="poisson", data=datasetCONW)
CONW.noYYYY = glmer(AB_CONW ~ State_P_500 + LOG_BCR_CONW + offset(OF_CONW) + (1|ClusterSP), family="poisson", data=datasetCONW)
```

### These models are working:
```{r modelsRun, echo=TRUE, include=TRUE}
CONW.noCl = glmer(AB_CONW ~ State_P_500 + LOG_BCR_CONW + offset(OF_CONW) + (1|YYYY), family="poisson", data=datasetCONW)
CONW.noClInt = glmer(AB_CONW ~ State_P_500 + LOG_BCR_CONW + offset(OF_CONW) + (1|YYYY) + (1|ClusterSP:YYYY), family="poisson", data=datasetCONW)
CONW.noRE = glm(AB_CONW ~ State_P_500 + LOG_BCR_CONW + offset(OF_CONW), family="poisson", data=datasetCONW)
```

As always, Eliot was right! I thought I had done it, but I had only dropped the interaction and YYYY, forgot to test dropping Cluster or all RE. My bad, sorry! :)

So, we have 3 options:
1. Model with one random effect as year: (1|YYYY)
```{r model1}
CONW.noCl
```

2. Model with one random effect as year and interation between year and cluster: (1|YYYY) + (1|ClusterSP:YYYY)
```{r model2}
CONW.noClInt
```

3. Model without random effects.
```{r model3}
CONW.noRE
```

The smallest IAC was for the option 2: 803.98.

