---
title: "SupMat_April2019"
author: "Tati Micheletti"
date: "12 April 2019"
output: pdf_document
---
 Table S1 : We ran three models at the local scale and four at the neighborhood scale for each of 15 
 songbird species. Sample size refers to the number of surveys used to evaluate each question. Random 
 effects and offsets were consistent in all models. “BCR” stands for “Bird Conservation Region”. “Jurisdiction” 
 indicates the political jurisdiction (i.e., Canadian Provinces and Territories). “LCC” stands for “Land Cover 
 Class”.
 
```{r S1, eval = TRUE, echo = FALSE, warning = FALSE, message=FALSE}
library("reproducible")
library("data.table")
dt <- data.table(prepInputs(url = "https://drive.google.com/open?id=13eHQmibKxOrP1OwGbQEVyqi90rsUzGeo",
                            destinationPath = getwd(), fun = "data.table::fread",
                            filename2 = NULL))

localBoth <- NROW(dt)
localUndisturbed <- NROW(dt[State_P_100 == 0])
localSuccesional <- NROW(dt[State_P_100 > 0 & Agent_L == "Transitional"]) + localUndisturbed
localAlienating <- NROW(dt[State_P_100 > 0 & Agent_L != "Transitional"]) + localUndisturbed

neighBoth <- NROW(dt)
neighUndisturbed <- NROW(dt[State_P_500 == 0])
neighSuccesional <- NROW(dt[State_P_500 > 0 & Agent_N == "Transitional"]) + neighUndisturbed
neighAlienating <- NROW(dt[State_P_500 > 0 & Agent_N != "Transitional"]) + neighUndisturbed
neighBothUnds <- NROW(dt[State_P_500 > 0 &
                           State_P_100 == 0]) + neighUndisturbed
neighAlienatingUnds <- NROW(dt[State_P_500 > 0 & Agent_N != "Transitional" &
                              State_P_100 == 0]) + neighUndisturbed
neighSuccesionalUnds <- NROW(dt[State_P_500 > 0 & Agent_N == "Transitional" &
                                  State_P_100 == 0]) + neighUndisturbed

SampleSize <- data.table(sampleSize = c(localBoth,
                                        localAlienating,
                                        localSuccesional,
                                        neighBoth,
                                        neighAlienating,
                                        neighSuccesional,
                                        neighBothUnds,
                                        neighAlienatingUnds,
                                        neighSuccesionalUnds))
footNote <- data.table(footNote = c(localUndisturbed,
                                    neighUndisturbed))
                                        
knitr::kable(SampleSize)
knitr::kable(footNote)

```
 
 Table S2: Avian surveys were grouped by cluster and by year of sampling. This table shows 
 the number of clusters, surveys, and whether they were disturbed or undisturbed at the local 
 and neighborhood scale for each year of sampling. We identified a total of 483 non-overlapping 
 clusters; however, as some clusters contained stations that were surveyed in multiple years, 
 the sum of clusters across years is 702.  
 
```{r S2, eval = TRUE, echo = FALSE, warning = FALSE, message=FALSE}
dt <- dt[order(YYYY)]
dt <- dt[, .("sumClu" = length(unique(ClusterSP)), 
              "sumSurv" = .N,
              "LocalDist" = sum(State_P_100 > 0),
              "LocalUndist" = sum(State_P_100 == 0),
              "NeighDist" = sum(State_P_500 > 0),
              "NeighUndist" = sum(State_P_500 == 0)) , by = YYYY]
knitr::kable(dt)
```
 
 