---
title: "hindcastingPosthoc"
author: "tmichele"
date: "22 May 2020"
output: pdf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hold")
```

# Overview

Provide an overview of what the module does / how to use the module.

Module documentation should be written so that others can use your module.
This is a template for module documentation, and should be changed to reflect your module.

## R Markdown

R Markdown syntax allows R code, outputs, and figures to be rendered in the documentation.

For help writing in R Markdown, see http://rmarkdown.rstudio.com/.

# Usage

```{r module_usage}

updateGithubPackages <- FALSE
if (updateGithubPackages){
  devtools::install_github("PredictiveEcology/reproducible@development")
  devtools::install_github("achubaty/amc@development")
  devtools::install_github("PredictiveEcology/pemisc@development")
  devtools::install_github("PredictiveEcology/map@development")
  devtools::install_github("PredictiveEcology/SpaDES.core@developement") # Updates SpaDES.tools and SpaDES.core quickPlot
  devtools::install_github("PredictiveEcology/SpaDES.tools@development") # Updates SpaDES.tools and SpaDES.core quickPlot
  devtools::install_github("PredictiveEcology/LandR@development") # Updates SpaDES.tools and SpaDES.core quickPlot
  devtools::install_github("PredictiveEcology/usefulFuns@development") # Updates LandR
}
  library("SpaDES")
  library("reproducible")
  library("raster")
  library("qs")
  library("usefulFuns")
  
  setPaths(modulePath = file.path(getwd(), "modules"),
           cachePath = file.path(getwd(), "cache"),
           inputPath = file.path(getwd(), "inputs"),
           outputPath = file.path(getwd(), "outputs/posthocAnalysis/bayesian/trial8"))
  getPaths() # shows where the 4 relevant paths are
  # invisible(sapply(X = list.files(file.path(getwd(), "functions"), full.names = TRUE), FUN = source))
  
  times <- list(start = 1985, end = 2011)
  
  posthocFolder <- file.path(getwd(),"outputs/posthocAnalysis")
  modules <- list("hindcastingPosthoc")
  maskedDensityRasFolder <- file.path(posthocFolder, "maskedDensityRas")
  
  rasterToMatch <- prepInputs(url = "https://drive.google.com/open?id=1c-rnifCpK-qU9fkg__f4RP0ueRUNjjiR",
                                      targetFile = "rasterToMatch_fromFocalRasters.qs",
                                      destinationPath = Paths$inputPath,
                                      fun = "qs::qread")
  # StudyArea needs to be masked to the Boreal forest
  boreal <- Cache(defineStudyArea, testArea = TRUE, specificTestArea = "boreal",
                  mapSubset = NULL, userTags = c("where:global", "function:defineStudyArea", 
                                                 "location:boreal")) ## rp is the whole North American boreal region
  boreal <- projectInputs(boreal, targetCRS = crs(rasterToMatch))

  # Exclude USA and other countries and Alpine from the boreal
  toDropEco <- c("B_ALPINE", "H_ALPINE")
  toDropCountry <- c("USA", "GREENLAND", "FRANCE")
  boreal <- boreal[!boreal$TYPE %in% toDropEco & !boreal$COUNTRY %in% toDropCountry,]

  studyArea <- Cache(prepInputs, url = "https://drive.google.com/open?id=1tgqn8FajD1iSj0aECONGhFzwInau0-q8",
                         targetFile = "NIR2016_MF.shp", archive = "NIR2016_MF.zip",
                         alsoExtract = "similar",
                         overwrite = TRUE, 
                         studyArea  = boreal,
                         omitArgs = c("overwrite"),
                         rasterToMatch = rasterToMatch, 
                         destinationPath = Paths$inputPath, 
                         userTags = c("objectName:studyArea",
                                      "obs: borealForestExtent",
                                      "obs2:OnlyCanadaNoAlpine")) #cacheId = "30bda38785cd89e6"
  
boreal <- Cache(cropInputs, boreal, studyArea = studyArea) # cacheId = "0f8a459b727333c6"
predictedDT <- qread(file.path(Paths$outputPath, "predictedDT.qs"))

objects <- list("studyArea" = studyArea,
                "rasterToMatch" = rasterToMatch,
                "predictedDT" = predictedDT)
inputs <- list()
outputs <- list()

parameters <- list(
  # "hindcastingPosthoc" = list(
  #   "" = 
  # )
)
bayes <- simInitAndSpades(times = times, 
                          params = parameters, 
                          paths = getPaths(),
                          modules = modules, 
                          objects = objects, 
                          debug = 2)
 
```

# Parameters

Provide a summary of user-visible parameters.

```{r moduleParams, echo = FALSE}
df_params <- moduleParams("hindcastingPosthoc", "..")
knitr::kable(df_params)
```

# Events

Describe what happens for each event type.

## Plotting

Write what is plotted.

## Saving

Write what is saved.

# Data dependencies

## Input data

How to obtain input data, and a description of the data required by the module.
If `sourceURL` is specified, `downloadData("hindcastingPosthoc", "path/to/modules/dir")` may be sufficient.

```{r moduleInputs, echo = FALSE}
df_inputs <- moduleInputs("hindcastingPosthoc", "..")
knitr::kable(df_inputs)
```

## Output data

Description of the module outputs.

```{r moduleOutputs, echo = FALSE}
df_outputs <- moduleOutputs("hindcastingPosthoc", "..")
knitr::kable(df_outputs)
```

# Links to other modules

Describe any anticipated linkages to other modules.
