---
title: "Exploring BLPW"
author: "Tati Micheletti"
date: "7/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why this file?

On 30th July, after running Alberto's models, I found out that BLPW was the only species not to converge when trying to fit glmer models. I had then to use glm (no random effects). It converged, but the results are a bit troubling.

This only works if you have the mySimOut objects with all the results from the `splitModelPlot` as well as the `glmerBirdModels` from the  `borealBirdsAndForestry` project.

Here we have the source data (with which we fitted the models), as well as Alberto's original data (that he used on the submitted version of the manuscript):
```{r loading_data}
localTrans <- mySimOut@.envir$data$localTransitional
dataCompare <- localTrans[, c("PKEY", "ClusterSP", "BLPW", "AB_BLPW")]
```

## Checking minimum and maximum values
**Note that:** 
- AB_sp: Alberto's abundance data
- sp: SQL abundance data
```{r minMax}
min(dataCompare$BLPW, na.rm = T)
min(dataCompare$AB_BLPW, na.rm = T)
max(dataCompare$BLPW, na.rm = T)
max(dataCompare$AB_BLPW, na.rm = T)
```

No differences here (i.e. no outliers)

## Checking summary
Let's see the mean and the number of NA's:
```{r summary}
summary(dataCompare, na.rm = T)
```
Nothing seems too out of the ordinary here. However, the SQL has more NA's. The number of NA's should be the same considering the application of the filtering (NA's stand for the clusters that didn't have any counts of a given species, while 0's are for the clusters that counted at least 1 bird of a given species in any given year, in a cluster)

## Checking differing rows
Which rows differ on these values?
```{r differences}
dataCompare$rowsDiffer <- dataCompare$BLPW == dataCompare$AB_BLPW
dataCompare$bothNA <- is.na(dataCompare$BLPW) & is.na(dataCompare$AB_BLPW)
toFilter <- which(dataCompare$rowsDiffer == FALSE & dataCompare$bothNA == FALSE |
                    is.na(dataCompare$rowsDiffer) & dataCompare$bothNA == FALSE)
diffData <- dataCompare[toFilter]
clustersToCheck <- unique(diffData[is.na(rowsDiffer) & bothNA == FALSE, ClusterSP])
```
clustersToCheck are the cluster ID's that did not match the 0's and NA's from Alberto's data and the SQL.

## Checking specific different clusters
According to Alberto's file (`Steps to create database.docx`): 

3.For  each  species,  identified  clusters  with  no  presence  (AB  =  0)  
    a.Removed  the  abundance  data  for  those  clusters  (i.e.  replaced  0  by  NA)

From what I understand, this means that the clusters that have no observations (ie. no presence), the values for abundance are NA, while for clusters that have at least one observation have 0's for abundance when the specific point has no observations.

```{r clusters}
listSums <- unlist(lapply(X = clustersToCheck, FUN = function(cluster){
assign(x = paste0("SQL_sum", cluster), sum(dataCompare[ClusterSP == cluster]$BLPW, na.rm = TRUE))
assign(x = paste0("Alberto_sum", cluster), sum(dataCompare[ClusterSP == cluster]$AB_BLPW, na.rm = TRUE))  
})
)
any(listSums > 0)
```
Here we see that for some clusters for BLPW (n = 14), Alberto attributed `0` when the value should be `NA` as the clusters that have no data should have `NA`, not `0`. 

When we reapeat this with BBWA:
```{r BBWA}
dataCompareBBWA <- localTrans[, c("PKEY", "ClusterSP", "BBWA", "AB_BBWA")]
min(dataCompareBBWA$BBWA, na.rm = T)
min(dataCompareBBWA$AB_BBWA, na.rm = T)
max(dataCompareBBWA$BBWA, na.rm = T)
max(dataCompareBBWA$AB_BBWA, na.rm = T)
summary(dataCompareBBWA, na.rm = T)
dataCompareBBWA$rowsDiffer <- dataCompareBBWA$BBWA == dataCompareBBWA$AB_BBWA
dataCompareBBWA$bothNA <- is.na(dataCompareBBWA$BBWA) & is.na(dataCompareBBWA$AB_BBWA)
toFilter <- which(dataCompareBBWA$rowsDiffer == FALSE & dataCompareBBWA$bothNA == FALSE |
                    is.na(dataCompareBBWA$rowsDiffer) & dataCompareBBWA$bothNA == FALSE)
diffData <- dataCompareBBWA[toFilter]
clustersToCheck <- unique(diffData[is.na(rowsDiffer) & bothNA == FALSE, ClusterSP])
listSums <- unlist(lapply(X = clustersToCheck, FUN = function(cluster){
assign(x = paste0("SQL_sum", cluster), sum(dataCompareBBWA[ClusterSP == cluster]$BBWA, na.rm = TRUE))
assign(x = paste0("Alberto_sum", cluster), sum(dataCompareBBWA[ClusterSP == cluster]$AB_BBWA, na.rm = TRUE))  
})
)
any(listSums > 0)
```

Here we have the same results for BBWA, where `0` was attributed to abundances of clusters that have no data and should be `NA`.