---
title: "Correlation between State_P_100 and State_P_500"
author: "Tati Micheletti"
date: "6/26/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Correlation test

I plotted and ran a correlation test on the dataset to determine how correlated `State_P_100` and `State_P_500` are.

First, I loaded the dataset:

```{r datasetLoading}
datasetFull <- data.table::fread(paste0("/home/tmichele/Documents/",
                                        "GitHub/borealBirdsAndForestry/modules/",
                                        "glmerBirdModels/data/Final_points_2010.csv"))
```

Then I got the subset of interest: 
```{r datasetSusetting}
subsetData <- data.frame(LOCAL = datasetFull$State_P_100, 
                         NEIGHBORHOOD = datasetFull$State_P_500)
```

## Plot

Then I plotted the values:  

```{r plot, echo = FALSE}
attach(subsetData)
plot(LOCAL, NEIGHBORHOOD)
```

## Correlation test

I then ran the correlation test:  

```{r dataCorrelation}
cor.test(LOCAL, NEIGHBORHOOD,
                        method = "pearson",
                        conf.level = 0.95)
```

##Conclusion

We conclude that `State_P_100` (Local disturbance %) and `State_P_500` (Neighborhood disturbance %) are highly correlated according to Pearson's correlation test.  
It presented a R of **\textcolor{red}{0.86}** and a significancy of less than **\textcolor{red}{0.01}**.




